---
title: "Fetching and Standardizing"
author: "Amira Swedan"
date: "`r Sys.Date()`"
output: word_document
---


## Overview

This data pipeline automates the process of RTM/mVAM data management for the RTM core variables related to the key indicators. The standard pipeline procedures are aligned with [RBC RTM SOPs](https://wfp-my.sharepoint.com/:w:/g/personal/amira_swedan_wfp_org/EVistHIsU_xAsdsF-2SEcRoB-cMrwS-IuyjyscsUVLfuig?e=SH9hHK) that includes a detailed explanation about the methodology applied in this script.

The key objective of the standard pipeline is to maintain the same process of data management across all RBC countries running RTM/mVAM activities.

[**The key functions implemented in this section are**:]{.underline}

-   Fetching the raw dataset
-   Renaming and selecting core variables


Before running this pipeline script, you need to update the input file included in the script dependencies *RTM_pipeline_input.xlsx*, Specify in this file the start and end date that you wish to extract the data within. [*Check readme for more details*]

The script running environment is included in the *readme* file, [make sure that you have all required packages installed]{.underline} before running the script


## Load packages and read relevant input files

```{r load packages, echo=TRUE, message=FALSE, warning=FALSE}

# load required packages

library(tidyverse)
library(readxl)
library(httr)
library(jsonlite)
library(lubridate)
library(lares)
library(labelled)
library(expss)
library(ggplot2)
library(ggrepel)
library(tools)
library(writexl)

# read input file

pipInput = read_excel('Pipeline input/RTM_pipeline_input.xlsx' , sheet = 'main_parameters')
stdNames = read_excel('Pipeline input/RTM_pipeline_input.xlsx' , sheet = 'standard_names_mapping')
pipInput$StartDate = format(as.Date(pipInput$StartDate), "%m/%d/%Y")
pipInput$EndDate = format(as.Date(pipInput$EndDate), "%m/%d/%Y")

```


## Connect and Extract Raw Data through Crystal API (Syria and Yemen)

Connect through Crystal API to fetch the raw data.

To connect you need to pass the key parameters (`APIKey`, `start date`, `end date`) for each country separately to *Crystal_API_connect* function.

The parameters should be inserted in the pipeline input file, the configuration list for each country (`Syria`, and `Yemen`) will be updated automatically.



The new variables names in this section are aligned with the standard variables names included in WFP codebook.

The codebook is embedded in WFP [survey designer](https://www.surveydesigner.vam.wfp.org/design/survey).


```{r pull raw data through Crystal API , message=FALSE, warning=FALSE}
##########################################################
#the function below could be used to pull the survey data for any country if the provider is Crystal
###########################################################

Crystal_API_connect <- function (con_config) {
  
json_body <- jsonlite::toJSON(con_config, auto_unbox = TRUE)

print (json_body)
  
 r =  POST(url = 'http://campaignapi.crystelcall.com/CampaignAPI/api/Campaign' , body = json_body ,  add_headers('Content-Type'='application/json')) 
 
 if (r$status_code == 200) {
   print ('Crystal API connection is ok ')
 }
 else {
     print (paste0('ERROR:reading from Crystal API returned the status code ' , r$status_code))
 }
 
 Rawdata = fromJSON(rawToChar(r$content))
 
 return(Rawdata)
 
}



## create a configuration list for countries utilizing Crystal API (Currently only Yemen and Syria)

Syria_config = list ('APIKey' = paste("Syria_" , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Syria'] , "/")) [1] , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Syria'] , "/")) [3] , sep = ""),  'DateFrom' = pipInput$StartDate[pipInput$CountryCode == 'Syria']  , 'DateTo' = pipInput$EndDate[pipInput$CountryCode == 'Syria'] )

Yemen_config = list ('APIKey' = paste("Yemen1_" , unlist (strsplit(pipInput$StartDate[pipInput$CountryCode == 'Yemen1'] , "/")) [1] , unlist (strsplit(pipInput$StartDate[pipInput$CountryCode == 'Yemen1'] , "/")) [3] , sep = ""),  'DateFrom' = pipInput$StartDate[pipInput$CountryCode == 'Yemen1']  , 'DateTo' = pipInput$EndDate[pipInput$CountryCode == 'Yemen1'] )

```

## Renaming

```{r}
##Variables standard names mapping

replace_variable_names <- function(data, old_names, new_names) {
  # Convert the data to a data frame if it's not already
  if (!is.data.frame(data)) {
    data <- as.data.frame(data)
  }
  
  # Check if the lengths of old_names and new_names are the same
  if (length(old_names) != length(new_names)) {
    stop("The lengths of old_names and new_names must be the same.")
  }
  
  # Iterate over the old_names and new_names to replace variable names
  for (i in seq_along(old_names)) {
    if (old_names[i] %in% names(data)) {
      names(data)[names(data) == old_names[i]] <- new_names[i]
    }
  }
  
  # Return the updated data frame
  return(data)
}

```


```{r}
##Pull the data for only countries indicated in the input list and export the raw data in csv file
####################################################

if ('Syria' %in% pipInput$CountryCode) {

SyriaRaw = Crystal_API_connect (Syria_config)
SyriaRaw = SyriaRaw[SyriaRaw$Completed == 'Y' ,]
print(paste0('total number of completed surveys for Syria survey is ' , nrow(SyriaRaw)))
SyriaRaw = replace_variable_names(SyriaRaw , stdNames$Orignial[stdNames$Country == 'Syria'] , stdNames$Standard[stdNames$Country == 'Syria'])
write.csv(SyriaRaw , paste0("Raw data/Syria/Syria_raw_" , pipInput$SvyID[pipInput$CountryCode == 'Syria'] , "_" , Sys.Date() ,  ".csv"))
print ( paste0('Syria ', pipInput$SvyID[pipInput$CountryCode == 'Syria'] , ' raw data is exported' ))

}


if ('Yemen1' %in% pipInput$CountryCode ) {
YemenRaw = Crystal_API_connect(Yemen_config)
YemenRaw = YemenRaw[YemenRaw$Completed == 'Y' ,]
print(paste0('total number of completed surveys for Yemen survey is ' , nrow(YemenRaw)))
names(YemenRaw) = gsub("[\r\n]", "", names(YemenRaw))
YemenRaw = replace_variable_names(YemenRaw , stdNames$Orignial[stdNames$Country == 'Yemen'] , stdNames$Standard[stdNames$Country == 'Yemen'])
write.csv(YemenRaw , paste0("Raw data/Yemen/Yemen_raw_" , pipInput$SvyID[pipInput$CountryCode == 'Yemen1'] ,"_"  , Sys.Date(), ".csv"))
print ( paste0('Yemen ', pipInput$SvyID[pipInput$CountryCode == 'Yemen1'] , ' raw data is exported' ))

}
```




## import raw data directly (Iraq, Gaza, Ukraine, Lebanon) 

For countries that utilizes MODA, there is a need to manually extract the data from MODA and include it in the country respective folder under the raw data folder. 
Make sure while extracting the data through MODA to:
1. export it in csv
2. don't include the groups names in the header
3. extract the data with the variables labels
4. include only the data files that are relevant to the process in the respective country folder(s) in raw data container 

```{r message=FALSE, warning=FALSE}

if ('Iraq' %in% pipInput$CountryCode) {
  setwd("Raw data//Iraq")
  files = list.files()
  IrqRaw = read.csv(files[1] , header = TRUE)
  IrqRaw = IrqRaw[(IrqRaw$CallDispo == "Someone answers"  & IrqRaw$RESPConsent == "Yes"),]
  print(paste0('total number of completed surveys for Iraq survey is ' , nrow(IrqRaw)))
  IrqRaw = replace_variable_names(IrqRaw , stdNames$Orignial[stdNames$Country == 'Iraq'] , stdNames$Standard[stdNames$Country == 'Iraq'])
}




```