---
title: "RTM standard pipeline processes"
author: "Amira Swedan"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This data pipeline automates the process of RTM/mVAM data management for the RTM core variables related to the key indicators. The standard pipeline procedures are aligned with [RBC RTM SOPs](https://wfp-my.sharepoint.com/:w:/g/personal/amira_swedan_wfp_org/EVistHIsU_xAsdsF-2SEcRoB-cMrwS-IuyjyscsUVLfuig?e=SH9hHK) that includes a detailed explanation about the methodology applied in this script.

The key objective of the standard pipeline is to maintain the same process of data management across all countries running RTM/mVAM activities.

[**The key functions implemented in this standard pipeline are**:]{.underline}

-   Fetching the raw dataset
-   Renaming and selecting core variables
-   Data quality checks
-   Data re-coding and merging
-   Indicators computation
-   Weights construction
-   Connecting to the regional master table

Before running this pipeline script, you need to update the input file included in the script dependencies *RTM_pipeline_input.xlsx*, Specify in this file the start and end date that you wish to extract the data within. [*Check readme for more details*]

In addition to modifying the input file, make sure that all population figures included in the weights input table file are updated, and the desired indicators are selected.

The script running environment is included in the *readme* file, [make sure that you have all required packages installed]{.underline} before running the script

## Load packages and read relevant input files

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(xlsx)
library(httr)
library(jsonlite)
library(lubridate)
library(lares)
library(labelled)
library(expss)


pipInput = read_excel('RTM_pipeline_input.xlsx' , sheet = 'main_parameters')
stdNames = read_excel('RTM_pipeline_input.xlsx' , sheet = 'standard_names_mapping')
pipInput$StartDate = format(as.Date(pipInput$StartDate), "%m/%d/%Y")
pipInput$EndDate = format(as.Date(pipInput$EndDate), "%m/%d/%Y")

```

## Connect and Extract Raw Data

Connect through Crystal API to fetch the raw data.

To connect you need to pass the key parameters (`APIKey`, `start date`, `end date`) for each country separately to *Crystal_API_connect* function.

The parameters should be inserted in the pipeline input file, the configuration list for each country (`Lebanon`,`Syria`, and `Yemen`) will be updated automatically.

To add a new country, first modify the input file and create a new configuration list that should be passed to the connection function.

```{r extract raw data, message=FALSE, warning=FALSE, echo=TRUE}

## create a configuration list for each country in the pipeline ## Don't proceed with running the pipeline functions for a country without creating this list 

Syria_config = list ('APIKey' = paste("Syria_" , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Syria'] , "/")) [1] , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Syria'] , "/")) [3] , sep = ""),  'DateFrom' = pipInput$StartDate[pipInput$CountryCode == 'Syria']  , 'DateTo' = pipInput$EndDate[pipInput$CountryCode == 'Syria'] )

Yemen_config = list ('APIKey' = paste("Yemen1_" , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Yemen1'] , "/")) [1] , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Yemen1'] , "/")) [3] , sep = ""),  'DateFrom' = pipInput$StartDate[pipInput$CountryCode == 'Yemen1']  , 'DateTo' = pipInput$EndDate[pipInput$CountryCode == 'Yemen1'] )

Leb_config = list ('APIKey' = "Lebanon_112022",  'DateFrom' = pipInput$StartDate[pipInput$CountryCode == 'Lebanon']  , 'DateTo' = pipInput$EndDate[pipInput$CountryCode == 'Lebanon'] )

Libya_config = list ('APIKey' = paste("Libya_" , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Libya'] , "/")) [1] , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Libya'] , "/")) [3] , sep = ""),  'DateFrom' = pipInput$StartDate[pipInput$CountryCode == 'Libya']  , 'DateTo' = pipInput$EndDate[pipInput$CountryCode == 'Libya'] )

##########################################################
#the function below could be used to pull the survey data for any country if the provider is Crystal
###########################################################

Crystal_API_connect <- function (con_config) {
  
json_body <- jsonlite::toJSON(con_config, auto_unbox = TRUE)

print (json_body)
  
 r =  POST(url = 'http://campaignapi.crystelcall.com/CampaignAPI/api/Campaign' , body = json_body , add_headers('Content-Type'='application/json')) 
 
 if (r$status_code == 200) {
   print ('Crystal API connection is ok ')
 }
 else {
     print (paste0('ERROR:reading from Crystal API returned the status code ' , r$status_code))
 }
 
 Rawdata = fromJSON(rawToChar(r$content))
 
 return(Rawdata)
 
}
######################################################
##Pull the data for only countries provided in the input list and export the raw data in csv file
####################################################

if ('Syria' %in% pipInput$CountryCode) {

SyriaRaw = Crystal_API_connect (Syria_config)
SyriaRaw = SyriaRaw[SyriaRaw$Completed == 'Y' ,]
print(paste0('total number of completed surveys for Syria survey is ' , nrow(SyriaRaw)))
write.csv(SyriaRaw , paste0("Syria_raw_" , pipInput$SvyID[pipInput$CountryCode == 'Syria'] , ".csv"))
print ( paste0('Syria ', pipInput$SvyID[pipInput$CountryCode == 'Syira'] , ' raw data is exported' ))
}


if ('Yemen1' %in% pipInput$CountryCode ) {
YemenRaw = Crystal_API_connect(Yemen_config)
YemenRaw = YemenRaw[YemenRaw$Completed == 'Y' ,]
print(paste0('total number of completed surveys for Yemen survey is ' , nrow(YemenRaw)))
write.csv(YemenRaw , paste0("Yemen_raw_" , pipInput$SvyID[pipInput$CountryCode == 'Yemen1'] , ".csv"))
}

if ('Lebanon' %in% pipInput$CountryCode) {
LebRaw = Crystal_API_connect(Leb_config)
LebRaw = LebRaw[LebRaw$Completed == 'Y' ,]
print(paste0('total number of completed surveys for Lebanon survey is ' , nrow(LebRaw)))
write.csv(LebRaw , paste0("Lebanon_raw_" , pipInput$SvyID[pipInput$CountryCode == 'Lebanon'] , ".csv"))
}


if ('Libya' %in% pipInput$SurveyType) {
LibyaRaw = Crystal_API_connect(Libya_config)
LibyaRaw = LibyaRaw[LibyaRaw$Completed == 'Y' ,]
print(paste0('total number of completed surveys for Libya survey is ' , nrow(LibyaRaw)))
write.csv(LebRaw , paste0("Libya_raw_" , pipInput$SvyID[pipInput$CountryCode == 'Libya'] , ".csv"))
}


```

## Renaming variables according to the standard codebook, and selecting only core indicators

The new variables names in this section are specified according to the standard variables names included in WFP codebook.

The codebook is embedded in WFP [survey designer](https://www.surveydesigner.vam.wfp.org/design/survey).

```{r renaming and selecting, echo=TRUE, message=FALSE, warning=FALSE}

##Variables standard names mapping

replace_variable_names <- function(data, old_names, new_names) {
  # Convert the data to a data frame if it's not already
  if (!is.data.frame(data)) {
    data <- as.data.frame(data)
  }
  
  # Check if the lengths of old_names and new_names are the same
  if (length(old_names) != length(new_names)) {
    stop("The lengths of old_names and new_names must be the same.")
  }
  
  # Iterate over the old_names and new_names to replace variable names
  for (i in seq_along(old_names)) {
    if (old_names[i] %in% names(data)) {
      names(data)[names(data) == old_names[i]] <- new_names[i]
    }
  }
  
  # Return the updated data frame
  return(data)
}



SyriaRaw = replace_variable_names(SyriaRaw , stdNames$Orignial[stdNames$Country == 'Syria'] , stdNames$Standard[stdNames$Country == 'Syria'])

YemenRaw = replace_variable_names(YemenRaw , stdNames$Orignial[stdNames$Country == 'Yemen'] , stdNames$Standard[stdNames$Country == 'Yemen'])

LebRaw = replace_variable_names(LebRaw , stdNames$Orignial[stdNames$Country == 'Lebanon'] , stdNames$Standard[stdNames$Country == 'Lebanon'])

# select core indicators - Syria

SyriaRaw_reduced = SyriaRaw %>% select (ADMIN0Name,ADMIN1Name, ADMIN2Name ,RspID,SvyDate, ObsDate ,EnuName,HHH_YN,HHHSex,HHHAge,HHHEducation,RESPAge,HHSize , HHDispl, IDPDuration,HDwellType,HWaterSRC, FCSStap, FCSPulse, FCSVeg, FCSFruit,FCSPr, FCSDairy, FCSFat , FCSSugar, FCS_SRf, MktAccess_1M,  rCSILessQlty, rCSIBorrow, rCSIMealNb , rCSIMealSize, rCSIMealAdult, Lcs_stress_DomAsset, Lcs_stress_Saving, Lcs_stress_SellFoodRation, Lcs_stress_CrdtFood, Lcs_crisis_OutSchool, Lcs_crisis_ProdAssets, Lcs_crisis_Migration, Lcs_em_ResAsset, LcsR_em_FemAnimal, Lcs_em_Begged)

#Yemen
YemenRaw_reduced = YemenRaw %>% select(RspID, SvyDate,ObsDate,EnuName,RESPAge , HHHSex, ADMIN0Name , ADMIN1Name, ADMIN2Name , HHSize, HHDispl, HHChronIllNb , FCSStap , FCSPulse, FCSVeg, FCSFruit, FCSPr, FCSDairy, FCSFat, FCSSugar , FCS_SRf , rCSILessQlty, rCSIBorrow , rCSIMealNb , rCSIMealSize , rCSIMealAdult ,LcsEN_stress_DomAsset , LcsEN_stress_CrdtFood , LcsEN_stress_Saving, LcsEN_stress_BorrowCash, LcsEN_crisis_ProdAssets, LcsEN_crisis_Health , LcsEN_crisis_OutSchool , LcsEN_em_ResAsset , LcsEN_em_Begged, LcsEN_em_LastAnimal , LhCSIEnAccess , HHIncFirst_SRi , HHIncChg_1M  )

#Lebanon
LebRaw_reduced = LebRaw %>%  select(EnuName, ObsDate, RespNationality , HHSizebelow2 , HHSizebetween2_5, HHSizebetween5_17, HHSizebetween18_59 , HHSize60above, HHHSex, HWaterSRC, HToiletType , HElectricitySRC , ADMIN0Name , ADMIN1Name , ADMIN2Name , SvyDate,RESPAge,HDwellType, FCSStap, FCSPulse, FCSDairy, FCSPr, FCSVeg, FCSFruit, FCSFat, FCSSugar, rCSILessQlty, rCSIBorrow, rCSIMealSize, rCSIMealNb, rCSIMealAdult = rCSIMealAdul, Lcs_stress_DomAsset ,  Lcs_stress_CrdtFood , Lcs_crisis_ProdAssets , Lcs_em_ResAsset, Lcs_stress_BorrowCash  , Lcs_stress_LessSchool , Lcs_crisis_OutSchool , Lcs_crisis_Health  , Lcs_em_Begged  , FCS_SRf  )

```

## Re-coding and Merging

```{r recoding, echo=TRUE, message=FALSE, warning=FALSE}

#Syria
SyriaRaw_reduced = SyriaRaw_reduced %>% mutate(HHHSex=recode(HHHSex, '1' ~ 'Female' , '2' ~ 'Male' ) , HHDispl = recode(HHDispl , '1' ~ 'Resident' , '2' ~ 'Returnee' , '3' ~ 'IDP') , HHHEducation = recode(HHHEducation , '1' ~ 'None' , '2' ~ 'Primary' , '3' ~ 'Secondary' , '4' ~ 'Vocational' , '5' ~ 'University') , IDPDuration = recode(IDPDuration, '1' ~ 'LT3M' , '2' ~'3To6M'  , '3' ~ '7To12M' , '4' ~ 'Above12M' ) , HDwellType = recode(HDwellType, '1' ~ 'Own' , '2' ~ 'Rent' , '3' ~ 'Guest_family' , '4' ~ 'Guest_strangers' , '5' ~ 'SharingAcco' , '6' ~ 'Hotel' ,'7' ~ 'RentAnotherH' , '8' ~'NoPlace' , '9' ~ 'TempShelter' , '10' ~ 'Camp' , '11' ~ 'UnfinishedShelter' , '12' ~ 'CollectiveShelter' , '13' ~ 'Other') , HWaterSRC = recode(HWaterSRC, '1' ~ 'Piped' , '2' ~ 'PublicTap' , '3' ~ 'borehole' ,'4' ~ 'ProtectedWell' , '5' ~ 'ProtectedSpring' , '6' ~ 'UnprotectedSpring' , '7' ~ 'River' , '8' ~ 'Lake' , '9' ~ 'Rain' , '10' ~ 'BottledWater' , '11' ~ 'TruckedWater'), FCS_SRf = recode(FCS_SRf, '1' ~ 'OwnProd' , '2' ~ 'Labour' , '3' ~ 'Purchase' , '4' ~ 'FoodAssist' , '5' ~ 'Scavenging' , '6' ~ 'Gift') , MktAccess_1M = recode (MktAccess_1M , '1' ~ 'Yes' , '2' ~ 'No' , '3' ~ 'DK') )

lcs_vars = names (SyriaRaw_reduced %>% select (starts_with('Lcs')))
SyriaRaw_reduced[lcs_vars] = replaceall( SyriaRaw_reduced [lcs_vars] , c('1' , '2' , '3' , '4') , c('Yes' , 'No' , 'AlreadyExhausted'  , 'NotApplicable') , quiet = TRUE)

#Yemen
YemenRaw_reduced = YemenRaw_reduced %>% mutate(HHHSex=recode(HHHSex, '1' ~ 'Male' , '2' ~ 'Female' ) , HHDispl = recode(HHDispl , '1' ~ 'IDP' , '2' ~ 'Resident' ), FCS_SRf = recode(FCS_SRf, '1' ~ 'OwnProd' , '2' ~ 'Hunting' , '3' ~ 'Purchase' , '4' ~ 'On_Credit' , '5' ~'Gift' ,  '6' ~ 'Begging' , '7' ~ 'Swap' , '8' ~ 'FoodAssist' , '9' ~ 'CasualLabour') , HHIncFirst_SRi = recode (HHIncFirst_SRi , '1' ~ 'RegEmp' , '2' ~ 'CasualLabour' , '3' ~ 'NoWork_assist') , HHIncChg_YN_1M = recode(HHIncChg_1M, '1' ~ 'Increased' ,'2' ~ 'Same' , '3' ~ 'Reduced' , '4' ~ 'Stopped'), LhCSIEnAccess = recode(LhCSIEnAccess, '1' ~ 'Food' , '2' ~ 'Education' , '3' ~ 'Health' , '4' ~ 'Shelter' , '5' ~ 'Sanitation' , '6' ~ 'Other') )

lcs_vars = names (YemenRaw_reduced %>% select (starts_with('Lcs')))
YemenRaw_reduced[lcs_vars] = replaceall( YemenRaw_reduced [lcs_vars] , c('1' , '2' , '3' , '4' , '99') , c('No' , 'AlreadyExhausted' , 'Yes'  , 'No_shortage' , 'NotApplicable') , quiet = TRUE)

#Lebanon
LebRaw_reduced = LebRaw_reduced %>% mutate(RespNationality=recode(RespNationality, '1' ~'Leb' , '2' ~ 'Syr_Leb' , '3' ~ 'Pal_Leb' , '4' ~ 'Syr' , '5' ~ 'Pal' , '6' ~ 'Irq' , '7' ~ 'Migrant' , '8' ~ 'Others' ) , HHHSex=recode(HHHSex, '1' ~ 'Male' , '2' ~ 'Female') , HWaterSRC = recode(HWaterSRC, '1' ~ 'Piped' , '2' ~ 'PublicTap' , '3' ~ 'Well_pump' ,'4' ~ 'ProtectedWell' , '5' ~ 'Distilled_water' , '6' ~ 'UnprotectedSpring' , '7' ~ 'tank_truck' , '8' ~ 'car_tank' , '9' ~ 'water_Seller' , '10' ~ 'bottled_water' , '11' ~ 'surface_water' , '12' ~ 'Other') , HToiletType = recode(HToiletType, '1' ~ 'Flush_toilet' , '2' ~ 'hole_tiledbath' , '3' ~ 'hole_unpavedbath' , '4' ~ 'bucket' , '5' ~ 'OpenAir') , HElectricitySRC = recode(HElectricitySRC , '1' ~ 'company' , '2' ~ 'NonOwned_Generator' , '3' ~ 'Owned_Generator' , '4' ~ 'Relatives' , '5' ~ 'Solar') , FCS_SRf = recode(FCS_SRf , '1' ~ 'OwnProd' , '2' ~ 'Purchase' ,'3' ~ 'Labour'  ,'4' ~ 'Gift' , '5' ~ 'FoodAssist' , '6' ~ 'Gov' , '7' ~ 'Community' , '8' ~ 'Other') )

lcs_vars = names (LebRaw_reduced %>% select (starts_with('Lcs')))
LebRaw_reduced[lcs_vars] = replaceall( LebRaw_reduced [lcs_vars] , c('1' , '2' , '3' , '4') , c('No' , 'AlreadyExhausted' , 'Yes'  , 'NotApplicable') , quiet = TRUE)


## Create the master table 

df_master = full_join(SyriaRaw_reduced , YemenRaw_reduced)
df_master = full_join( df_master, LebRaw_reduced)

```

## Food Security Indicators Calculation

```{r indicators , echo=TRUE, message=FALSE, warning=FALSE}


#assign variable and value labels
var_label(df_master$FCSStap) <- "Cereals, grains, roots and tubers"
var_label(df_master$FCSPulse) <- "Pulses/ legumes / nuts"
var_label(df_master$FCSDairy) <- "Milk and other dairy products"
var_label(df_master$FCSPr) <- "Meat, fish and eggs"
var_label(df_master$FCSVeg) <- "Vegetables and leave"
var_label(df_master$FCSFruit) <- "Fruits"
var_label(df_master$FCSFat) <- "Oil/fat/butter"

#calculate FCS and FCG

FCS_vars = names (df_master %>% select (starts_with('FCS')))

df_master[FCS_vars[-9]] = apply(df_master[FCS_vars[-9]] , 2 , as.numeric)


df_master <- df_master %>% mutate(FCS = (2 * FCSStap) +(3 * FCSPulse) +(4*FCSPr) +(4*FCSDairy) + FCSVeg  + FCSFruit +(0.5*FCSFat) +(0.5*FCSSugar))
var_label(df_master$FCS) <- "Food Consumption Score"

df_master <- df_master %>% mutate(FCG = case_when(
  FCS <= 28 ~ 1, between(FCS, 28.5, 42) ~ 2, FCS > 42 ~ 3))

val_lab(df_master$FCG) = num_lab("
             1 Poor
             2 Borderline
             3 Acceptable
")

var_label(df_master$FCG) <- "FCS Categories"

#rCSI (reduced Coping Strategies Index)
#assign variable and value labels
var_label(df_master$rCSILessQlty) <-  "Rely on less preferred and less expensive food in the past 7 days"
var_label(df_master$rCSIBorrow) <- "Borrow food or rely on help from a relative or friend in the past 7 days"
var_label(df_master$rCSIMealNb) <-  "Reduce number of meals eaten in a day in the past 7 days"
var_label(df_master$rCSIMealSize) <- "Limit portion size of meals at meal times in the past 7 days"
var_label(df_master$rCSIMealAdult) <-  "Restrict consumption by adults in order for small children to eat in the past 7 days"

#calculate reduced Coping Strategy Index (rCSI)
rcsi_vars = names (df_master %>% select (starts_with('rCSI')))
df_master[rcsi_vars] = apply(df_master[rcsi_vars] , 2 , as.numeric)

df_master <- df_master %>% mutate(rCSI = rCSILessQlty + (2 * rCSIBorrow) + rCSIMealNb + rCSIMealSize + (3 * rCSIMealAdult))
var_label(df_master$rCSI) <- "Reduced coping strategies index (rCSI)"



```

## Quality Checks and Associated Flags

```{r calculate livelihood coping indicators, echo=TRUE, message=FALSE, warning=FALSE}
df_master

```


## Regional Master Table Connection and Feeding

```{r upload data, echo=TRUE, message=FALSE, warning=FALSE}

```
