---
title: "RTM standard pipeline processes"
author: "Amira Swedan"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This data pipeline automate the process of RTM/mVAM data management for the RTM core variables related to the key indicators. The standard pipeline procedures are aligned with [RBC RTM SOPs](https://wfp-my.sharepoint.com/:w:/g/personal/amira_swedan_wfp_org/EVistHIsU_xAsdsF-2SEcRoB-cMrwS-IuyjyscsUVLfuig?e=SH9hHK) that includes a detailed explanation about the methodology applied in this script.

The key objective of the standard pipeline is to maintain the same process of data management across all countries running RTM/mVAM activities.

[**The key functions implemented in this standard pipeline are**:]{.underline}

-   Fetching the raw dataset
-   Renaming and selecting core variables
-   Data quality checks
-   Data re-coding and merging
-   Indicators computation
-   Weights construction
-   Connecting to the regional master table

Before running this pipeline script, you need to update the input file included in the script dependencies *RTM_pipeline_input.xlsx*, Specify in this file the start and end date that you wish to extract the data within. see the table below for more details:

+------------------------------------------------+-------------------------------------------------------------------------------------+-------------------------------------------------------------------------+-------------------------------------------------------+----------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+
| **Provider**                                   | **SurveyType**                                                                      | **Country Code**                                                        | **StartDate**                                         | **EndDate**                                        | **SurveyID**                                                                                                                                 |
+------------------------------------------------+-------------------------------------------------------------------------------------+-------------------------------------------------------------------------+-------------------------------------------------------+----------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+
| include here the name of the service provider. | Specify if the survey is RTM or other type of surveys ( include only RTM or Others) | the country code is the name of the country as included in the API key. | the first date the you need to extract the data from. | the last date that you need to extract the data to | This is unique ID to differentiate each data collection round. It is sequential for each country. and should be unique across all countries. |
|                                                |                                                                                     |                                                                         |                                                       |                                                    |                                                                                                                                              |
|                                                |                                                                                     |                                                                         | It should be in this format                           | It should be in this format                        |                                                                                                                                              |
|                                                |                                                                                     |                                                                         |                                                       |                                                    |                                                                                                                                              |
|                                                |                                                                                     |                                                                         | MM/DD/YYYY                                            | MM/DD/YYYY                                         |                                                                                                                                              |
+------------------------------------------------+-------------------------------------------------------------------------------------+-------------------------------------------------------------------------+-------------------------------------------------------+----------------------------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+

: Input table structure

In addition to modifying the input file, make sure that all population figures included in the weights input table file are updated, and the desired indicators are selected.

The script running environment is included in the *readme* file, [make sure that you have all required packages installed]{.underline} before running the script

## Load packages and read relevant input files

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(httr)
library(jsonlite)
library(lubridate)

pipInput = read_excel('RTM_pipeline_input.xlsx')
pipInput$StartDate = format(as.Date(pipInput$StartDate), "%m/%d/%Y")
pipInput$EndDate = format(as.Date(pipInput$EndDate), "%m/%d/%Y")

```

## Connect and extract raw data

Connect through Crystal API to fetch the raw data.

To connect you need to pass the key parameters (`APIKey`, `start date`, `end date`) for each country separately to *Crystal_API_connect* function.

The parameters should be inserted in the pipeline input file, the configuration list for each country (`Lebanon`,`Syria`, and `Yemen`) will be updated automatically.

To add a new country, first modify the input file and create a new configuration list that should be passed to the connection function.

```{r extract raw data, message=FALSE, warning=FALSE, echo=TRUE}


Syria_config = list ('APIKey' = paste("Syria_" , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Syria'] , "/")) [1] , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Syria'] , "/")) [3] , sep = ""),  'DateFrom' = pipInput$StartDate[pipInput$CountryCode == 'Syria']  , 'DateTo' = pipInput$EndDate[pipInput$CountryCode == 'Syria'] )

Yemen_config = list ('APIKey' = paste("Yemen1_" , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Yemen1'] , "/")) [1] , unlist (strsplit(pipInput$EndDate[pipInput$CountryCode == 'Yemen1'] , "/")) [3] , sep = ""),  'DateFrom' = pipInput$StartDate[pipInput$CountryCode == 'Yemen1']  , 'DateTo' = pipInput$EndDate[pipInput$CountryCode == 'Yemen1'] )

Leb_config = list ('APIKey' = "Lebanon_112022",  'DateFrom' = pipInput$StartDate[pipInput$CountryCode == 'Lebanon']  , 'DateTo' = pipInput$EndDate[pipInput$CountryCode == 'Lebanon'] )

##########################################################
Crystal_API_connect <- function (con_config) {
  
json_body <- jsonlite::toJSON(con_config, auto_unbox = TRUE)

print (json_body)
  
 r =  POST(url = 'http://campaignapi.crystelcall.com/CampaignAPI/api/Campaign' , body = json_body , add_headers('Content-Type'='application/json')) 
 
 if (r$status_code == 200) {
   print ('Crystal API connection is ok ')
 }
 else {
     print (paste0('ERROR:reading from Crystal API returned the status code ' , r$status_code))
 }
 
 Rawdata = fromJSON(rawToChar(r$content))
 
 return(Rawdata)
 
}
######################################################

SyriaRaw = Crystal_API_connect (Syria_config)
SyriaRaw = SyriaRaw[SyriaRaw$Completed == 'Y' ,]
print(paste0('total number of completed surveys for Syria survey is ' , nrow(SyriaRaw)))

YemenRaw = Crystal_API_connect(Yemen_config)
YemenRaw = YemenRaw[YemenRaw$Completed == 'Y' ,]
print(paste0('total number of completed surveys for Yemen survey is ' , nrow(YemenRaw)))

LebRaw = Crystal_API_connect(Leb_config)
LebRaw = LebRaw[LebRaw$Completed == 'Y' ,]
print(paste0('total number of completed surveys for Lebanon survey is ' , nrow(LebRaw)))

######################################################
## Export the raw data
#####################################################


write.csv(SyriaRaw , paste0("Syria_raw_" , pipInput$SvyID[pipInput$CountryCode == 'Syria'] , ".csv"))

write.csv(YemenRaw , paste0("Yemen_raw_" , pipInput$SvyID[pipInput$CountryCode == 'Yemen1'] , ".csv"))

write.csv(LebRaw , paste0("Lebanon_raw_" , pipInput$SvyID[pipInput$CountryCode == 'Lebanon'] , ".csv"))

```

```         
```

## Renaming variables according to the standard codebook, and selecting only core indicators

The new variables names in this section are specified according to the standard variables names included in WFP codebook.

The codebook is embedded in WFP [survey designer](https://www.surveydesigner.vam.wfp.org/design/survey).

```{r renaming and selecting, echo=TRUE, message=FALSE, warning=FALSE}
# Syria
SyriaRaw_reduced = SyriaRaw %>% select (ADMIN0Name = Country, ADMIN1Name = City, ADMIN2Name = Area,RspID,SvyDate, ObsDate= LogDate1 ,EnuName = Agent ,HHH_YN,HHHSex,HHHAge,HHHEducation,RESPAge,HHSize , HHDispl, IDPDuration,HDwellType =HouseType,HWaterSRC, FCSStap, FCSPulse, FCSVeg, FCSFruit,FCSPr, FCSDairy, FCS_SRf, MktAccess_1M, MktAccessCnstr, rCSILessQlty, rCSIBorrow, rCSIMealNb , rCSIMealSize, rCSIMealAdult, Lcs_stress_DomAsset, Lcs_stress_Saving, Lcs_stress_SellFoodRation, Lcs_stress_CrdtFood, Lcs_crisis_OutSchool, Lcs_crisis_ProdAssets, Lcs_crisis_Migration, Lcs_em_ResAsset, LcsR_em_FemAnimal, Lcs_em_Begged)

#Yemen
YemenRaw_reduced = YemenRaw %>% select(RspID, SvyDate,ObsDate= LogDate1 ,EnuName = Agent ,RESPAge = BirthYear, HHHSex = HoHSex, ADMIN0Name = Country, ADMIN1Name = City, ADMIN2Name = Area, HHSize= HHSize_90days, HHDispl = IDP_YN, HHChronIllNb = HH65_Disease, FCSStap = Staples, FCSPulse = Pulses, FCSVeg = Veg, FCSFruit = Fruits, FCSPr = Proteins, FCSDairy = Dairy, FCSFat = Fats, FCSSugar = Sugars, FCS_SRf = FoodSrc1, rCSILessQlty = LessExpensiveFood, rCSIBorrow = BorrowOrHelp, rCSIMealNb = ReduceNumMeals, rCSIMealSize = LimitPortionSize, rCSIMealAdult = RestrictConsumption,LcsEN_stress_DomAsset = LC_2_SoldAsset, LcsEN_stress_CrdtFood = LC_3_CreditFood, LcsEN_stress_Saving = LC_4_SpentSaving, LcsEN_stress_BorrowCash = LC_5_BorrowMoney, LcsEN_crisis_ProdAssets = LC_6_SoldProduction, LcsEN_crisis_Health = LC_7_ReduceHealth, LcsEN_crisis_OutSchool = LC_8_WithdrawSchool, LcsEN_em_ResAsset = LC_9_SoldHome, LcsEN_em_Begged = LC_10_Begged , LcsEN_em_LastAnimal = LC_11_LastAnimal, LhCSIEnAccess = LC_MainReason, HHIncFirst_SRi = Income_MainSrc, HHIncChg_YN_1M = Income_Change  )

#Lebanon
LebRaw_reduced = LebRaw %>%  select(EnuName = Agent, ObsDate= LogDate1, Relation_to_HoH , DispStat_Nationality , HHSizebelow2 , HHSizebetween2_5, HHSizebetween5_17, HHSizebetween18_59 , HHSize60above, HHHSex = HoHH_Sex, HWaterSRC= HHdrinkingwater, HToiletType = HHtoiletfacility, HElectricitySRC =MainSourcesElectricity, ADMIN0Name = Country, ADMIN1Name = City, ADMIN2Name = Area, SvyDate,RESPAge,HDwellType= HHHousingCurrent, FCSStap, FCSPulse, FCSDairy, FCSPr, FCSVeg, FCSFruit, FCSFat, FCSSugar, rCSILessQlty, rCSIBorrow, rCSIMealSize, rCSIMealNb, rCSIMealAdult = rCSIMealAdul, Lcs_stress_DomAsset =LhCSIDomAsset,  Lcs_stress_CrdtFood = LhCSICrdtFood, Lcs_crisis_ProdAssets = LhCSIProdAsset, Lcs_em_ResAsset= LhCSIsoldhouse, Lcs_stress_BorrowCash = LhCSIBorrowCash , Lcs_stress_LessSchool = LhCSIchangedschool, Lcs_crisis_OutSchool = LhCSIOutSchool , Lcs_crisis_Health = LhCSIHealth , Lcs_em_Begged = LhCSIBegged , FCS_SRf = HHFoodSource, HHIncFirst_SRi= Sources_of_income1 , HHIncSec_SRi= Sources_of_income2 , HHIncThird_SRi=Sources_of_income3)

```

## Re-coding and merging

```{r recoding, echo=TRUE, message=FALSE, warning=FALSE}

#Syria
SyriaRaw_reduced = SyriaRaw_reduced %>% mutate(HHHSex=recode(HHHSex, '1'='Female' , '2' = 'Male' ) , HHDispl = recode(HHDispl , '1' = 'Resident' , '2' = 'Returnee' , '3' = 'IDP') , HHHEducation = recode(HHHEducation , '1' = 'None' , '2' = 'Primary' , '3' = 'Secondary' , '4' = 'Vocational' , '5' = 'University') ,  )





```

## Quality checks and flags

```{r Check data quality, echo=TRUE, message=FALSE, warning=FALSE}

```

## Food security indicators calculation

```{r calculate livelihood coping indicators, echo=TRUE, message=FALSE, warning=FALSE}

```

## Construct weights

```{r Construct weights, echo=TRUE, message=FALSE, warning=FALSE}

```

## Connect to regional master table

```{r upload data, echo=TRUE, message=FALSE, warning=FALSE}

```
